What's in a Name? \\ Linear Temporal Logic \\ Literally Represents Time Lines

### 2023-08-28 paper/main.tex
clarified fixes...
added FIX label
clarify

%KYR: FIX THIS!
The algorithm is outlined below (see Algorithms 1 - 4). %\cref{alg:ba2wregex}

@@

\begin{definition}[Timeline Length]
    We define the timeline length of a regular expression $\tllen(A)$ recursively as follows.
    \begin{align*}
        \tllen(\emptyset)   &= -\infty\\
        \tllen(\epsilon)        &= 0\\
        \tllen(a)           &= 1\\
        \tllen(A_1 A_2)     &= \tllen(A_1) + \tllen(A_2)\\
        \tllen(A_1 + A_2)   &= \max(\tllen(A_1), \tllen(A_2))\\
        \tllen(A^*)         &= \tllen(A)
    \end{align*}
    We also extend this definition to $\omega$-regular expressions as follows.
    \begin{align*}
        \tllen(A^\omega)    &= \tllen(A)\\
        \tllen(AB)          &= \tllen(A) + \tllen(B)\\
        \tllen(B_1 + B_2)   &= \max(\tllen(B_1), \tllen(B_2))
    \end{align*}
    Then $\tllen$ captures the length of the longest path in our timeline graph visualization.
\end{definition}

@@

%KYR: Here we nicely show (with pictures as much as possible) all of the individual statistics and results we gathered (note aggregation occurs below).
%\paragraph{Real-world datasets}
We study the feasibility of our tool on a benchmark suite of LTL formulas gathered from real world use cases.
%
We gather a test suite totalling 91 formulas from two real-world requirement specification applications. We take $6$ specifications from NASA's Automated Airspace Concept (AAC)~\cite{GCMTR16}. We also collect formulas from the Acacia suite of examples~\cite{RV10}. We reap $85$ formulas from the suite of $23$ examples, by generating separate visualizations for each line of specification. We choose to visualize each formula individually, rather than visualize a complete set of (conjuncted) specifications, as a complete specification set may describe a large, complex system, and we only target the use case of validating individual formulas. % We choose to visualize each formula specifically, rather than complete specifications, as

We run \texttt{ltl2regex} on these $91$ examples: $87$ of the $91$ input formulas complete execution within a timeout of $20$s. Within the $87$, two examples have regular expressions of star height $8$ and are intractable to generate graph visualizations of. We are able to generate tractable visualization on the remaining $85$, or over $93\%$ of benchmarks tested. The plots of visualization complexity on these formula are provided below in \cref{fig:real-world}.

@@

%\paragraph{Scalability}
{\bf Scalability.}
%KYR: FIX!!!
We measure the scalability of our tool by running it against a set of scalable LTL formulas. We define a scalable LTL formula as one that describes an $n$-bit number using a binary counter, as generated by~\cite{RV10}. As $n$ gets larger, the LTL formula required to encode the $n$-bit counter scales in size. \cref{fig:scalability} shows that the timeline length grows approximately exponentially with the number of bits in the counter, which matches the result in Rozier and Vardi~\cite{RV10}.

In \cref{fig:scalability}, the $x$-axis ranges from $1$ to $6$, after which point it hits the $30$ seconds timeout we set, as the timeline length grows exponentially fast.

%KYR: FIX: is timeout 20 seconds or 30 seconds?!?!?

@@

\section{Artifact Availability: ``Thought is free\textsuperscript{\textsection}" }
\begingroup\renewcommand\thefootnote{\textsection}
\footnotetext{William Shakespeare, \emph{The Tempest}}
\endgroup
The tool we present in this paper is available at
\[
    \texttt{\url{https://github.com/EULIR/ltl-explainability}}
\]
which comes with two command-line tools, \texttt{ltl2regex} and \tool. The specific usage of the tool can be found  \href{https://github.com/EULIR/ltl-explainability#usage}{here}.

>>>>

%KYR: FIX THIS! We need to make it clear the relationship between Algorithms 1..4 and the figure displaying the high-level algorithm.
The algorithm is outlined below (see Algorithms 1 - 4). %\cref{alg:ba2wregex}

@@

\begin{definition}[Timeline Length]
    We define the timeline length of a regular expression $\tllen(A)$ recursively as follows.
    \begin{align*}
        \tllen(\emptyset)   &= -\infty\\  %KYR:FIX!!!
        \tllen(\epsilon)        &= 0\\
        \tllen(a)           &= 1\\
        \tllen(A_1 A_2)     &= \tllen(A_1) + \tllen(A_2)\\
        \tllen(A_1 + A_2)   &= \max(\tllen(A_1), \tllen(A_2))\\
        \tllen(A^*)         &= \tllen(A)
    \end{align*}
    We also extend this definition to $\omega$-regular expressions as follows.
    \begin{align*}
        \tllen(A^\omega)    &= \tllen(A)\\
        \tllen(AB)          &= \tllen(A) + \tllen(B)\\
        \tllen(B_1 + B_2)   &= \max(\tllen(B_1), \tllen(B_2))
    \end{align*}
    Then $\tllen$ captures the length of the longest path in our timeline graph visualization.
\end{definition}

@@

%KYR: Here we nicely show (with pictures as much as possible) all of the individual statistics and results we gathered (note aggregation occurs below).
%\paragraph{Real-world datasets}
{\bf Feasibility.}
We study the feasibility of our tool on a benchmark suite of LTL formulas gathered from real world use cases.
%
We gather a test suite totalling 91 formulas from two real-world requirement specification applications. We take $6$ specifications from NASA's Automated Airspace Concept (AAC)~\cite{GCMTR16}. We also collect formulas from the Acacia suite of examples~\cite{RV10}. We reap $85$ formulas from the suite of $23$ examples, by generating separate visualizations for each line of specification. We choose to visualize each formula individually, rather than visualize a complete set of (conjuncted) specifications, as a complete specification set may describe a large, complex system, and we only target the use case of validating individual formulas. % We choose to visualize each formula specifically, rather than complete specifications, as

We run \texttt{ltl2regex} on these $91$ examples: $87$ of the $91$ input formulas complete execution within a timeout of $20$ seconds for the feasibility tests. Within the $87$, two examples have regular expressions of star height $8$ and are intractable to generate graph visualizations of. We are able to generate tractable visualization on the remaining $85$, or over $93\%$ of benchmarks tested. The plots of visualization complexity on these formula are provided below in \cref{fig:real-world}.

@@

%\paragraph{Scalability}
{\bf Scalability.}
%KYR: FIX!!! The description of the exact experiment run is not clear and seems to contradict the points in the above paragraph...
We measure the scalability of our tool by running it against a set of \emph{scalable} LTL formulas. We define a scalable LTL formula as one that describes an $n$-bit number using a binary counter, as generated by~\cite{RV10}. As $n$ gets larger, the LTL formula required to encode the $n$-bit counter scales in size. \cref{fig:scalability} shows that the timeline length grows approximately exponentially with the number of bits in the counter, which matches the result in Rozier and Vardi~\cite{RV10}.

In \cref{fig:scalability}, the $x$-axis ranges from $1$ to $6$, after which point it hits the $30$ seconds timeout we set for the scalability tests, as the timeline length grows exponentially fast.

%KYR: FIX: is timeout 20 seconds or 30 seconds?!?!? It still sounds like this is the same test as the paragraph above but with a different timeout. Please describe the difference in the tests; what was run exactly for a 30 second timeout and why was this different than the 20 second timout?

@@

\section{Artifact Availability: ``Thought is free\textsuperscript{\textsection}'' }
\begingroup\renewcommand\thefootnote{\textsection}
\footnotetext{William Shakespeare, \emph{The Tempest}}
\endgroup
The tool we present in this paper is available at
\[
    \texttt{\url{https://github.com/EULIR/ltl-explainability}}
\]
which comes with two command-line tools, \texttt{ltl2regex} and \tool. The specific usage of the tool can be found  \href{https://github.com/EULIR/ltl-explainability#usage}{here}.
